% !TeX root = vpl.tex

\chap{Counting (Advanced)}\label{ch.counting}

In this chapter we show how states of the Thymio robot can be used to
count numbers and even perform simple arithmetic.

The design and implementation of the projects will not be presented
in detail. We assume that you have enough experience by now to develop
them yourself. The source code of working programs is included in the
archive, but don't look at them unless you really have difficulties
solving a problem.

These projects use the \emph{clap} event to change states and
the default behaviour of the circle LEDs to display the state:

\importantbox{The current state of the robot is displayed
in the circle LEDs on the top of the robot.
\cref{fig.state-leds} shows the robot in the state
\bu{(on, on, on, on)}.}

Feel free to change either of these behaviours.

\sect{Odd and even}

\begin{quote}
\textbf{Program}\\Choose one of the quarters of the state.
It will be \bu{off} (white) if the number of claps
is even and \bu{on} (orange) if the number of claps is odd.
Touching the center button will reset to even
(since zero is an even number).
\end{quote}

{\raggedleft \hfill Program file \bu{count-to-two.aesl}}

Our method of counting demonstrates the concept of
\emph{modulo 2 arithmetic}.
We count starting from 0 to 1 and then back to 0.
The term \emph{modulo} is like the term \emph{remainder}:
if there have been 7 claps, then dividing 7 by 2 gives 3 and
remainder 1.
We only keep the remainder 1.

In modulo 2 arithmetic, 0 and 1 are often called even and odd,
respectively.

Another term for the same concept is \emph{cyclic arithmetic}.
Instead of counting from 0 to 1 and then from 1 to 2,
we \emph{cycle} back to the beginning:
0, 1, 0, 1, \ldots.

These concepts are very familiar because they are used in clocks.
Minutes and seconds are computed modulo 60 and hours are computed modulo
12 or 24. Thus, the second after 59 is not 60; instead, we cycle around
and start counting from 0 again. Similarly, the hour after 23 is not 24,
but 0. If the time is 23:00 and we agree to meet after 3 hours,
then the time set for the meeting is 26 modulo 24, which is 02:00 in the
morning.

\sect{Counting in unary}

\begin{quote}
Modify the program to count modulo 4.
There are four possible remainders, 0, 1, 2, 3.
Choose three quarters, one each to
represent the values 1, 2 and 3; the value 0 will be represented
by setting all quarters to \bu{off}.
\end{quote}

This method of representing numbers is called
\emph{unary representation}
because different elements of a state represent different numbers.
We often use unary representation to keep track of the count
of some objects; for example,
\begin{picture}(35,10)
\multiput(5,0)(5,0){4}{\put(0,0){\line(0,1){10}}}
\put(0,0){\line(3,1){25}}
\put(32,0){\line(0,1){10}}
\end{picture}
represents 6.

{\raggedleft \hfill Program file \bu{count-to-four.aesl}}

\exercisebox{\thechapter.1}{How high can we count on the Thymio using
unary representation?}

\sect{Counting in binary}

We are very familiar with \emph{based representation},
in particular base 10 (decimal) representation.
The symbols 256 in base 10 representation don't represent three
different objects.
Instead, the 6 represents the number of 1's, the 5 represents the
number of 10$\times$1=10's, and the 2 represents the number of
10$\times$10$\times$1=100's. Adding these factors gives the number
two hundred and fifty-six. Using base 10 representation,
we can write very large
numbers in a compact representation.
Furthermore, arithmetic on large numbers
is relatively easy using the methods we learned at school.

We use base 10 representation because we have 10 fingers
so it is easy to learn to use the representation.
Computers, however, have two ``fingers'' (\bu{off} and \bu{on})
so base 2 arithmetic is used in
computation. Base 2 arithmetic looks
strange at first; while we use the familiar symbols 0 and 1
also used in base 10, the rules for counting are cyclic at 2
instead of cyclic at 10:
\begin{displaymath}
0, 1, 10, 11, 100, 101, 110, 111, 1000, \ldots
\end{displaymath}

Given a base 2 number such as 1101, we compute its value from right
to left just as in base 10.
The rightmost digit represents the number of 1's, the next digit
represents the number of 1$\times$2=2's, the third digit represents
the number of 1$\times$2$\times$2=4's, and the leftmost digit
represents the
numbers of 1$\times$2$\times$2$\times$2=8's.
Therefore, 1101 represents 1+0+4+8, which is thirteen,
represented in base 10 as 13.

\newpage

\begin{quote}
\textbf{Program}\\
Modify the program for counting modulo 4 to use binary representation
\end{quote}

{\raggedleft \hfill Program file \bu{count-to-four-binary.aesl}}

We only need two quarters of the state to represent the numbers
0--3 in base 2.
Let the upper right quarter represent the number of 1's,
\bu{off} (white) for none and \bu{on} (orange) for one,
and let the upper left quarter represent the number of 2's.
For example, \blksm{state-right} represents the number 1 and
\blksm{state-left} represents the number 2.
If both quarters are white, the state represents 0, and if both
quarters are orange, the state represents 3.

There are four transitions $0\rightarrow 1, 1\rightarrow 2, 2
\rightarrow 3, 3\rightarrow 0$, so four event-action pairs
are needed, in addition to a pair to reset the program
when the center button is touched.

\medskip

\trickbox{The two bottom quarters are not used, so they are left gray
and are ignored by the program.}

\medskip

\exercisebox{\thechapter.2}{Extend the program so that it counts
modulo 8. The lower left quarter will represent the number of 4's.}

{\raggedleft \hfill Program file \bu{count-to-eight.aesl}}

\exercisebox{\thechapter.3}{How high can we count on the Thymio using
binary representation?}

\sect{Adding and subtracting}

Writing the program to count to 8 is quite tedious because you had to
program 8 event-action pairs, one for each transition from $n$ to $n+1$
(modulo 8). Of course, that is not how we count in a based
representation;
instead, we have methods for performing addition by adding the digits
in each place and carrying to the next place. In base 10
representation:
\begin{displaymath}
\begin{array}{r}
387\\
+426\\
\rule[1pt]{1.5em}{1pt}\\
813\\
\end{array}
\end{displaymath}
and similarly in base 2 notation:
\begin{displaymath}
\begin{array}{r}
0011\\
+1011\\
\rule[1pt]{2em}{1pt}\\
1110\\
\end{array}
\end{displaymath}
When adding 1 to 1, instead of 2, we get 10.
The 0 is written in the same column and we carry the 1 to the next
column to the left.
The example above shows the addition of 3 (=0011) and 11 (=1011)
to obtain 14 (=1110).

\begin{quote}
\textbf{Program}\\
Write a program that starts with a representation of 0.
Each clap adds 1 to the number.
The addition is modulo 16, so adding 1 to 15 results in 0.
\end{quote}

Guidance:

\begin{itemize}
\item The lower right quarter will be used to represent the number of 8's.
\item If the upper right quarter representing the number
of 1's shows 0 (white),
simply change it to 1 (orange). Do this regardless of what the other
quarters show.
\item If the upper right quarter representing the number of 1's
shows 1 (orange),
change it to 0 (white) and then carry the 1.
There will be three event-action pairs,
depending on the location of the \emph{next} quarter showing 0
(white).
\item If all quarters show 1 (orange), the value of 15 is represented.
Adding 1 to 15 modulo 16 results in 0, represented by all quarters
showing 0 (white).
\end{itemize}


{\raggedleft \hfill Program file \bu{addition.aesl}}

\exercisebox{\thechapter.4}{Modify the program so that it starts
with the value 15 and subtracts one at each clap down to zero,
and then cyclically back to 15.}

{\raggedleft \hfill Program file \bu{subtraction.aesl}}

\exercisebox{\thechapter.5}{Place a sequence of short segments of
black tape on a light surface (or white tape on a dark surface).
Write a program that causes the Thymio to move forward and stop
when it has detected the fourth tape.}

This exercise is not easy: the strips of tape have to be
sufficiently wide so that the robot detects them,
but not so wide that more than one event occurs per strip.
You will also have to experiment with the speed of the robot.

{\raggedleft \hfill Program file \bu{count-tapes-four.aesl}}
