% !TeX root = vpl.tex

\chap{Project: Reading Barcodes}\label{ch.barcode}

Barcodes are universally used in supermarkets and elsewhere to identify
objects. The identification is a number or a sequence of symbols that is
different for each type of object. The identification is used to access
a database containing information about an object, such as its price.
Let use build a barcode reader from the Thymio robot.

\textbf{Specification}

\begin{enumerate}
\item Carefully measure the distance between two front horizontal
sensors and the width of a single sensor.
Using a piece of flexible cardboard, black tape and strips of aluminum
foil, contruct an object with various arrangements of the strips of foil:

\begin{center}
\gr{barcode}{.6}
\end{center}

\item Each configuration of the three center horizontal sensors will
represent a different code. (How many codes can there be?) For some or
all of these codes, implement event-action pairs that cause the top LEDs
to display a color, depending on the code identified.

\end{enumerate}

\textbf{Guidance}:

Since we are only using the three center sensors, the squares for the
outer sensors will be gray. For the center sensors, squares where the
reflecting foil must appear in the code are red, while squares where the
foil must not appear are white. For example, the following event-action
pair turns the top LEDs yellow for the code \texttt{on-off-on}:

\blkc{barcode1-3}

\bigskip

\textbf{Program file}: \bu{barcode.aesl}\\
This program identifies barcodes with foil opposite any two of the three
center sensors, as well as the code with no foil. You can extend it for
other configurations. 
