% !TeX root = vpl.tex

\chap{Event and Action Blocks}\label{a.blocks}

\section*{Event blocks}

\blksm{event-buttons} \textbf{Buttons}. Click on one or more of the images
of the buttons; they will turn red. An event will occur if the red
buttons are touched.

\bigskip\bigskip\bigskip\bigskip


\blksm{event-prox} \textbf{Horizontal sensors} (five at the front of the
Thymio and two at the back). Click on one or more of the small squares.
If the square is gray, the reading of the is ignored; if the square is
white, an event occurs if a lot of light is reflected; if the square is
black, an event occurs if little light is reflected.

\bigskip

\blksm{event-ground} \textbf{Bottom sensors} (two on the bottom of the
Thymio). Use like the block for the horizontal sensors.

\bigskip\bigskip

\blksm{event-prox-advanced}, \blksm{event-prox-ground-advanced}
\textbf{Sensors, advanced mode}. Use like the previous blocks. In
addition, the top slider sets the threshold above which an object is
detected and the bottom slider sets the threshold below which the
absence of an object is detected.

\bigskip\bigskip\bigskip\bigskip\bigskip\bigskip

\blksm{event-tap} \textbf{Tap}. An event occurs when the Thymio is
tapped.

\bigskip\bigskip

\blksm{event-tap-advanced} \textbf{Tap, advanced mode}. Use like the
previous block. Click on the small center or right circle to change to
an accelerometer event.

\bigskip\bigskip\bigskip\bigskip

\blksm{event-acc-roll}, \blksm{event-acc-pitch} \textbf{Accelerometer,
advanced mode}. Drag the red segment of the half-circle left or right.
An event will occur if the left / right or forwards / backwards angle
(respectively) of the Thymio is within the red segment.

\bigskip\bigskip\bigskip\bigskip\bigskip

\blksm{event-timer} \textbf{Timer, advanced mode}. An event occurs when
a timer has counted down to zero. The timer must have been set by a
previous timer \emph{action}.

\bigskip\bigskip\bigskip\bigskip

\blksm{event-state} \textbf{Event state, advanced mode}. The event
occurs only if the components of the current state match the
corresponding orange and white quarters of this block. The components
corresponding to the gray quarters need not match.

\bigskip\bigskip\bigskip\bigskip

\section*{Action blocks}

\blksm{action-motors} \textbf{Motors}. Move the left and right sliders
up to increase the forwards rotation of the left and right motors,
respectively, and move the sliders down to increase the backwards
rotation.

\bigskip\bigskip

\blksm{action-colors-up} \textbf{Top lights}. Move the three sliders to
the right to increase the red, green and blue components of the top
light, respectively. For example, moving the red and green sliders fully
to the right without moving the blue slider will turn the top light
yellow.

\bigskip\bigskip

\blksm{action-colors-down} \textbf{Bottom lights}. Turns the bottom
lights on. Use like the previous block.

\bigskip\bigskip\bigskip

\blksm{action-music} \textbf{Music}. The six small circles are notes. A
black note is short and a white note is long. Click on a circle to
change the length. The five horizontal bars, represent tones. Click on
the bar to move a circle to that bar.


\bigskip\bigskip\bigskip

\blksm{action-timer} \textbf{Timer, advanced mode} The timer can be set
for up to four seconds. Click anywhere within the white circle showing
the face of the clock. There will be a short animation and then the
amount of time until the alarm will be colored blue.

\bigskip\bigskip

\blksm{action-states} \textbf{States, advanced mode} The four
quarters in the block correspond to four components of a state.
Click a quarter to turn it to gray, orange or white.
